Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-07-24 exp1/iswc2013.tex
improvements to DoE explanation
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@622 43ef1190-90fd-4dd7-a769-d0121f02b80e

After measurements are done, we seek to estimate \emph{effects}. To
see how the \emph{main effects} can be estimated, consider a factor
$A$. We proceed to find the averages $a_1$ and $a_2$ of all measurements at level~1 and
on level~2 respectively. Then the main effect of $A$ is $a_2 - a_1$.

Similarly, \emph{interaction effects} for 2 factors $AB$ are defined
by comparing averages for equal versus non-equal levels of $A$ and
$B$.

This can be generalized to larger combinations of factors, but in
practice, this analysis is done by fitting a linear model to the data
with all factors and their interactions. It is shown in
\cite{wu2009experiments} that the effect is twice the coefficients of
the fit.

The next step is to understand which factors are important or
significant.  One approach is to plot the sorted effects against the
normal distribution. To understand why, note that if there is nothing
of interest, the measurements have a certain normal distribution
purely due to noise. When plotted against a normal distribution, the
plot would be a straight line. If there are any deviations from the
straight line, it implies that they may be significant. Such plots can
be found in
figures~\ref{fig:fullnormal},~\ref{fig:frac32normal},~\ref{fig:frac64normal}~and~\ref{fig:frac64hugenormal}. The
range in the $y$-axis is dictated by the number of runs.

@@

The experimentation is implemented in R\cite{rmanual}, which is a free
software environment for statistical computing and graphics. Necessary
tools for DoE has been implemented by the R community in packages
called DoE.base\cite{doebase} and FrF2\cite{frf2}. The experiments are
run on a third computer in a 1 gigabit Ethernet LAN with the two
experiment hosts. On the experiment hosts, four 4store instances runs
sequentially and independently on different ports, where processes
that are not actively answering a query are idle. Practically, the
``Machine'' factor specifies a hostname, whereas ``TripleC'' and
``Implement'' specifies a port number. On each host,

>>>>

We describe the influence of the choice of levels on the response in
terms of \emph{effects}.  For a factor $A$ with two levels, we let
$a_1$ and $a_2$ be the average response of all $2^{n-1}$ measurements
with $A$ at level 1, and 2, respectively.  The main effect of $A$ is
then defined as $a_2 - a_1$.

Similarly, \emph{interaction effects} for 2 factors $AB$ are defined
by comparing averages for equal versus non-equal levels of $A$ and
$B$.  Details of this theory, and how to compute the effects in
practice, using linear regression, are found in
\cite{wu2009experiments}.

The next step is to understand which factors are important or
significant.  One approach is to plot the sorted effects against the
normal distribution. To understand why, note that if there is nothing
of interest, the measurements have a certain normal distribution
purely due to noise. When plotted against a normal distribution, the
plot would be a straight line. If there are any deviations from the
straight line, it implies that they may be significant. Such plots can
be found in
Figs.~\ref{fig:fullnormal} to~\ref{fig:frac64hugenormal}. The
range in the $y$-axis is dictated by the number of runs.

@@

The experimentation is implemented in R\cite{rmanual}, which is a free
software environment for statistical computing and graphics. Necessary
tools for DoE has been implemented by the R community in packages
called DoE.base\cite{doebase} and FrF2\cite{frf2}. The experiments are
run on a third computer in a 1 gigabit Ethernet LAN with the two
experiment hosts. On the experiment hosts, four 4store instances runs
sequentially and independently on different ports, where processes
that are not actively answering a query are idle. Practically, the
``Machine'' factor specifies a hostname, whereas ``TripleC'' and
``Implement'' specifies a port number.
